A Survey of Exoskeleton Control Architectures and Algorithms\\ (Draft 3.0)

### 2015-09-28 exos/bleex.tex
cga-fixes
Added bibliographies by section to survey.tex. Use biber to build survey.tex
edits to survey document

\subsection{BLEEX}
\label{exo:bleex}

@@

Assuming no outside disturbances, the BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as $d$ (assuming no outside disturbances).   Neglecting gravity, the exokeleton angular velocity is modeled as 
\[v = G r + S d ,\] 
where $v$ represents the angular velocity of the exokeleton, $G$ is the transfer function from actuator inputs ($G$ is the exoskeleton's dynamics), $r$ is the actuator input, and $S$ is the sensitivity or transfer function from human torque to exokeleton angular velocity.

@@

The goal is to maximize sensitivity to $d$ \emph{without direct measurement.}  Sensitivity amplification accomplishes this by creating a feedback loop from a controller, $C$, acting only on exoskeleton variables.  A new sensitivity equation,
\[S_{new} = \frac{v}{d} = \frac{S}{1 + G C} ,\]
is maximized by applying positive feedback.  To achieve a large sensitivity, BLEEX uses $C = (1-\alpha^{-1})G^{-1}$ so that $\alpha$ provides a direct (scalar) amplification factor. A low pass filter is often added to the $C$ term in order to damp out high frequency dynamics of the exoskeleton, which are not captured in these models.  Note that the controller depends on an inverse dynamic model of the exoskeleton, $G^{-1}$.  Since the model is hybrid, these dynamics switch according to gait phases (single support, double support, stance).  BLEEX detects these transitions using foot sensors.  Assuming the single leg support dynamics are in the form,
\begin{equation}
\mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} + \mV({\bf\theta}) = {\bf T} + {\bf d} ,
\end{equation}
where ${\bf T}$ is a vector of actuator torques, the BLEEX control torque would be
\begin{equation}
{\bf T} = \mV({\bf\theta}) + (1 - \alpha^{-1}) \big [ \mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} \big ] .
\end{equation}
The user must provide the first torque component as the exoskeleton has no actuator acting between the foot and the ground (the system is underactuated).

@@

\bibliographystyle{plain}
\bibliography{exos/bleex}

@@

>>>>

\section{BLEEX}
\label{exo:bleex}
\begin{refsection}[exos/bleex.bib]

@@

The BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as $d$ (assuming no outside disturbances).   Neglecting gravity, the exokeleton angular velocity is modeled as 
\[v = G r + S d ,\] 
where $G$ is the transfer function from actuator inputs ($G$ is the exoskeleton's dynamics), $r$ is the actuator input, and $S$ is the sensitivity or transfer function from human torque to exokeleton angular velocity.

@@

The goal is to maximize sensitivity to $d$ \emph{without direct measurement.}  Sensitivity amplification accomplishes this by creating a feedback loop from a controller, $C$, acting only on exoskeleton variables.  A new sensitivity equation,
\[S_{new} = \frac{v}{d} = \frac{S}{1 + G C} ,\]
is maximized by applying positive feedback.  To achieve a large sensitivity, BLEEX uses $C = (1-\alpha^{-1})G^{-1}$ so that $\alpha$ provides a direct (scalar) amplification factor. A low-pass filter is often added to the $C$ term in order to damp out high frequency dynamics of the exoskeleton, which are not captured in these models.  Note that the controller depends on an inverse dynamic model of the exoskeleton, $G^{-1}$.  Since the model is hybrid, these dynamics switch according to gait phases (single support, double support, stance).  BLEEX detects these transitions using foot sensors.  Assuming the single leg support dynamics are in the form,
\begin{equation}
\mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} + \mV({\bf\theta}) = {\bf T} + {\bf d} ,
\end{equation}
where ${\bf T}$ is a vector of actuator torques, the BLEEX control torque would be
\begin{equation}
{\bf T} = \mV({\bf\theta}) + (1 - \alpha^{-1}) \big [ \mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} \big ] .
\end{equation}
The user must provide the first torque component as the exoskeleton has no actuator acting between the foot and the ground (the system is underactuated).

@@

\printbibliography[heading=subbibliography]

@@

\end{refsection}

### 2015-09-28 exos/exosText_MIT.tex
edits to survey document

Design is based  on fact that evidence from biology and passive walking devices suggest that legged locomotion can be very efficient.  The main physical concept behind this efficiency being that there is a gait cycle for a pair of legs that naturally exchanges energy between gravity and inertia.  The main mechanical design principles centered around using a combination of passive and active elements.  In particular, the MIT design uses {\bf springs, variable impedance joints, and powered actuators.}  The suit generally  functions as a lightweight, underactuated robot which runs in parallel to an operator and supports the weight of a payload.

Additionally, the leg structure of the suit allowed weight from the suit as well as payload to be transferred directly to the ground.

Additionally, there are several other design principles that the MIT team discusses which are especially relevant to this report.  The first is that joint {\bf powers scale linearly with mass}  The second is that altering the gait pattern of the operator has been shown to increase the physiological energy expended during locomotion.  Thus, the specifications for actuation and control for their system are extracted from the angle, torque , and power data of human walking joint patterns.

@@

The MIT exoskeleton uses a single actuator located at the hip.  The actuator is designed around a total system weight of 165 kg.  The maximum scaled hip torque for a system with this mass is approximately negative 130 Nm during the stance phase and approximately 100 Nm during the swing phase of the gait.  The hip is chosen as the actuated joint because proximal mass is metabolically less expensive in walking than distal mass.

The specific choice of actuation for the MIT exoskeleton hip joint was a {\bf linear series elastic actuator}.  This choice was made because it provided a lightweight and relatively inexpensive means of implementing force control with a bandwidth similar to that of natural muscle.  A model of the actuator is shown in Figure \ref{fig:mitSEA}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/mitSEA}
  \caption{}
  \vspace{-0.2in}
 \label{fig:mitSEA}   
 \end{figure} 
Like all series elastic designs the actuator shown in Figure \ref{fig:mitSEA} includes a spring in series with the output of the motor drive system of the actuator.  The motor-drive mechanism consists of a {\bf brushed DC motor driving a 3mm ball screw via a 2:1 reduction belt drive}.  The nut of the ball screw is coupled to the actuator output shaft via four compression die springs.  The compression of the die springs are measured using a linear potentiometer, which is how forces at the output are measured.

The motor chosen for the {\bf actuator was a 150W Maxon RE40 Brushed DC motor}, which was chosen based on the bio-mechanical data and for its power and weight ratio.

The force bandwidth for the actuator was characterized during both stance and swing phases of the walking gait cycle.  {\bf During stance, the bandwidth of the actuator was 35 Hz, while the bandwidth during swing was found to be 40Hz.}

@@

 The MIT exoskeleton design has a total of fourteen degrees of freedom, three at each hip, one at each knee, two at each ankle, and one in each foot.

 A cam mechanism was implemented at the hip joint to enable hip ab/adduction.  This system was necessary because, during abduction, there is was a length difference between the operator and exoskeleton leg which resulted from having non-collocated centers of rotation.  The cam mechanism automatically adjusts the exoskeleton leg length such that the center of rotation of the exoskeleton hip is projected onto the biological hip center of rotation.  A picture of the MIT exoskeleton is shown in Figure \ref{fig:MITsuit}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/MITsuit}
  \caption{}
%  \vspace{-0.2in}
 \label{fig:MITsuit}   
 \end{figure}

 The interface between the suit and operator consisted of shoulder straps, a compliant waist belt, thigh cuffs, and a shoe connection.

 The {\bf variable damper knee mechanism} in the flexion/extension degree of freedom was implemented using a magnetorheological damper.  The damper could exert a {\bf maximum braking torque of 60 Nm and consumed on average 1W of electrical power}.

 A linear spring located at the ankle was designed to capture negative energy during dorsilflexion.  The energy stored was released during plantar flexion to add energy to the system as the foot comes off the ground. A urethane spring was used as a liner spring in a lever compression assembly.  The spring constant of the urethane was 247 kN/m, the lever arm of the assembly was 0.0381 m, which gave the rotary spring a stiffness of 356 Nm/rad.

 A unidirectional spring assembly was also used in the hip ab/adduction degree of freedom.  The spring allowed the hip to freely abduct.  It was calculated that 8 Nm were required during adduction.  To this end, the effective rotary stiffness of the spring assembly was 96 Nm/rad.

The suit was powered using a 48V battery pack.

The suit was instrumented with rotary potentiometers at the hip and knee.  Additionally, strain gauges were placed on the structure of the exoskeleton shank, which measured sagittal bending moment and compression force

@@

 The control of the MIT exoskeleton functions around the fact that desired actuation and active damping at the knee are functions of gait cycle, at that the values of these desired functions could be determined from human walking data.  Fundamentally, the control strategy works as a state machine, which uses joint angle and measured forces to implement state transitions.

\begin{figure}[thpb]
\centering
\includegraphics[width=4.in]{exos/figs/MIT/walkingCycle}
  \caption{}
  \vspace{-0.2in}
 \label{fig:walkingCycle} 
 \end{figure} 
The walking cycle around which the MIT control scheme was designed is shown in Figure \ref{fig:walkingCycle}.  The individual control strategies in the various phases of the walking cycle are highlighted in Figure \ref{fig:walkingCycle}.  For the hip there are four distinct strategies.  During the thrust phase, the hip exerts a torque which helps to raise the center of mass of the system.  During the extension spring phase, a virtual spring stiffness is programmed which allows energy to be virtually stored while the center of mass moves forward.  During the swing assist phase the virtual energy is released, resulting in a torque being applied which assists in swinging the leg forward.  Finally, in leg retraction a torque is applied to help with foot placement and weight acceptance.  A diagram which illustrates how the sensors in the shank and hip joint are used to switch between these various phases is show in Figure \ref{fig:hipControl} 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/hipControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hipControl} 
 \end{figure}

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remained in the knee joint after it was turned off, creating a resistive torque.  The knee needed to be actively demagnetized at full extension during the last phase of stance to allow the knee to swing freely during the swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's onboard sensors is shown in Figure \ref{fig:kneeControl} 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 The MIT hardware and control system present a very well designed and thought out system which focus primarily on efficiency in design.  The series elastic linear ball screw hardware design seems to be a very smart way to implement efficient drives as well as to perform high fidelity force control.  Using a strain gauge in the shank was also a novel means of of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators as foot fall patterns for different users may include significant variations.

@@

>>>>

The MIT Exoskeleton is based on evidence from biology and passive walking devices that suggest that legged locomotion can be very efficient.  The main physical concept behind this efficiency being that there is a gait cycle for a pair of legs that naturally exchanges energy between gravity and inertia.  The main mechanical design principles center around the use of a combination of passive and active elements.  In particular, the MIT design uses {\bf springs, variable impedance joints, and powered actuators.}  The suit generally  functions as a lightweight, underactuated robot that runs in parallel to an operator and supports the weight of a payload. Additionally, the leg structure of the suit allows weight from the suit and payload to be transferred directly to the ground.

Of the several other design principles discussed, the MIT team emphasizes that joint {\bf powers scale linearly with mass}.  Also, the team points out that alterations in the operator's gait pattern have been shown to increase the physiological energy expended during locomotion.  Thus, the specifications for actuation and control for their system are extracted from the angle, torque , and power data of human walking joint patterns.

@@

The MIT exoskeleton uses a single actuator located at the hip.  The actuator is designed around a total system weight of 165 kg.  The maximum scaled hip torque for a system with this mass is approximately $-130$ Nm during the stance phase and approximately 100 Nm during the swing phase of the gait.  The hip is chosen as the actuated joint because proximal mass is metabolically less expensive in walking than distal mass.

The specific choice of actuation for the MIT exoskeleton hip joint was a {\bf linear series elastic actuator}.  This choice provided a lightweight and relatively inexpensive means of implementing force control with a bandwidth similar to that of natural muscle.  A model of the actuator is shown in Figure \ref{fig:mitSEA}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/mitSEA}
  \caption{}
  \vspace{-0.2in}
 \label{fig:mitSEA}   
 \end{figure} 
Like all series elastic designs, the actuator shown in Figure \ref{fig:mitSEA} includes a spring in series with the output of the motor drive system of the actuator.  The motor-drive mechanism consists of a {\bf brushed DC motor driving a 3mm ball screw via a 2:1 reduction belt drive}.  The nut of the ball screw is coupled to the actuator output shaft via four compression die springs.  The compression of the die springs are measured using a linear potentiometer, which is how forces at the output are measured.
The team selected a {\bf 150W Maxon RE40 Brushed DC motor} for the actuator  based on the bio-mechanical data and its power and weight ratio. The force bandwidth for this actuator was characterized during both stance and swing phases of the walking gait cycle.  {\bf During stance, the bandwidth of the actuator was 35 Hz, while the bandwidth during swing was found to be 40Hz.}

@@

 The MIT exoskeleton design has a total of fourteen degrees of freedom, three at each hip, one at each knee, two at each ankle, and one in each foot.  A cam mechanism implemented at the hip joint enables hip ab/adduction.  This system was necessary because, during abduction, there is a length difference between the operator and exoskeleton leg which results from having non-collocated centers of rotation.  The cam mechanism automatically adjusts the exoskeleton leg length such that the center of rotation of the exoskeleton hip is projected onto the biological hip center of rotation.  A picture of the MIT exoskeleton is shown in Figure \ref{fig:MITsuit}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/MITsuit}
  \caption{}
%  \vspace{-0.2in}
 \label{fig:MITsuit}   
 \end{figure} 
% 
 The interface between the suit and operator consists of shoulder straps, a compliant waist belt, thigh cuffs, and a shoe connection.

 The exoskeleton implements a {\bf variable damper knee mechanism} in the flexion/extension degree of freedom using a magnetorheological damper.  The damper can exert a {\bf maximum braking torque of 60 Nm and consums on average 1W of electrical power}.  Additionally, a linear spring at the ankles captures negative energy during dorsilflexion.  The energy stored is released during plantar flexion to add energy to the system as the foot comes off the ground. A urethane spring (247 kN/m) is used as a liner spring in a lever compression assembly.  With a lever arm of 0.0381 m, the assembly yields a spring rotary stiffness of 356 Nm/rad.

 A unidirectional spring assembly is also used in the hip ab/adduction degree of freedom.  The spring allows the hip to freely abduct.  The effective rotary stiffness of the spring assembly is 96 Nm/rad in order to provide the 8 Nm required during adduction.

The suit uses a 48V battery pack and is instrumented with \textbf{rotary potentiometers} at the hip and knee.  Additionally, \textbf{strain gauges} were placed on the structure of the exoskeleton shank to measured sagittal bending moment and compression force.

@@

The MIT exoskeleton controllers use the fact that desired actuation and active damping at the knee are functions of gait cycle.  Hence, the values of these desired functions can be determined from human walking data.  Fundamentally, the control strategy works as a state machine that uses joint angle and measured forces to implement state transitions.

\begin{figure}[thpb]
\centering
\includegraphics[width=4.in]{exos/figs/MIT/walkingCycle}
  \caption{}
  \vspace{-0.2in}
 \label{fig:walkingCycle} 
 \end{figure} 
The walking cycle around which the MIT control scheme was designed is shown in Figure \ref{fig:walkingCycle}.  The individual control strategies in the various phases of the walking cycle are highlighted in Figure \ref{fig:walkingCycle}.  For the hip there are four distinct strategies.  During the thrust phase, the hip exerts a torque that helps to raise the center of mass of the system.  During the extension spring phase, a virtual spring stiffness allows energy to be virtually stored while the center of mass moves forward.  During the swing assist phase the virtual energy is released, resulting in a torque being applied which assists in swinging the leg forward.  Finally, in leg retraction a torque is applied to help with foot placement and weight acceptance.  A diagram that illustrates how the sensors in the shank and hip joint are used to switch between these various phases is show in Figure \ref{fig:hipControl} 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/hipControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hipControl} 
 \end{figure}

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remains in the knee joint after it is turned off, creating a resistive torque.  The knee therefore needs to be actively demagnetized at full extension during the last phase of stance to allow it to swing freely during the subsequent swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's on-board sensors is in Figure \ref{fig:kneeControl} 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 The MIT hardware and control system present a very well designed and thought-out system that focuses primarily on efficiency in design.  The series elastic linear ball screw hardware design seems to be a very smart way to implement efficient drives as well as to perform high fidelity force control.  The use of a strain gauge in the shank was also a novel means of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators as foot fall patterns for different users may include significant variations.

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../survey"
%%% End:

### 2015-09-28 exos/exosText_dualMode.tex
edits to survey document

Exoskeleton suit was originally developed for military uses, specifically to allow operators to perform motions without incurring muscle fatigue.  Their main contributions in the literature re based on a {\bf virtual joint torque control}, which has been shown to be an effective human robot interaction control strategy in situations where the connections between the uses and hardware system are relatively unknown or can dramatically change as a function of time.

Additionally, the hardware system focuses on actuator selection.  The main hardware design principle is focused on developing a suit that can augment a users power and move fast when necessary.  They selected hydraulic actuation because the addition of electric actuator weight on the system leg joints makes the system difficult to move quickly.  On the other hand, hydraulic actuators have high power density relative to weight and size, and generally provide very high force outputs and low impedance.  The tradeoff with hydraulics come in the way of having low accuracy in terms of force control as well as hard to model nonlinear characteristics.

@@

The actuator specifications for the Hydraulic Lower Extremity Exoskeleton were designed around human walking data at 4km/h with a 45 kg load.  The resulting calculated hydraulic capacity for the while system was 2050psi and 8lpm.  The individual actuators were designed to be double acting at the hip and knee flexion/extension joints. The {\bf hip maximum thrust was 4 kN} and {\bf knee maximum thrust was 4 kN}.  Each joint joint included three-way servo valves (M 200, Star-Hydraulic) to control rate and direction of bi-directionality and bypass valves to connect the pump path to tank path for fast, energy efficient motion during swing phases.

@@

 The Hydraulic Lower Extremity Exoskeleton includes four actively controlled joints in the sagittal plane; the knee and hip flexion/extension degrees of freedom are powered degrees of freedom for each leg.  The hardware also includes eight passive joints: one ab/adduction DOF for each hip and three degrees of freedom for each ankle.

 The hydraulic system of the suit is composed of the four actuators and a hydraulic power unit, consisting of a pump, motor, temperature sensor, and various valves.

 Additionally, various {\bf sensors} are incorporated to measure joint angles, actuator forces , and ground reaction forces.  Three absolute angular encodes are incorporated at the hip, knee, and ankle pitch joints.  In-line load cells are incorporated in each cylinder tube to measure actuator forces, and four single axis load cells are included in foot sensors to be used for gait phase detection.

@@

 \begin{figure}[thpb]
\centering
\includegraphics[width=5.in]{exos/figs/hydLowerExrem/dualModeDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:dualModeDia}   
 \end{figure}
The control design principle for the Hydraulic Lower Extremity Exoskeleton is base around the concept that, for walking gaits, the {\bf stance and swing controllers should be considered separately}.  To this end, the literature for this hardware system introduces an active-passive control method called {\bf Dual-Mode Control}.  Active control is used during stance phases whereas the passive control is used to quickly and freely control the system's legs during swing phases.  Additionally, to handle the inherently discontinuous jumps in the command signals that occur as a function of this dual mode abstraction, a smoothing method is introduced that is used in conjunction with a pre-transition method that solves the swing delay due to internal cylinder pressure.  A diagram which highlights the different modes in this control strategy is shown in Figure \ref{fig:dualModeDia}.

Various components of the overall dual-mode control strategy, at the hardware level, are presented in Figure \ref{fig:suitDia}.  
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/suitDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:suitDia}   
 \end{figure}
This figure contains a number of components, specifically a {\bf gait phase estimation block} which uses the foot sensors to estimate the stance mode of the system, a {\bf human intention block}, which uses angular encoders and joint force sensors to estimate the motion of the operator, a {\bf torque controller}, and a {\bf gravity compensation} block.  Note that these various components of the control system are used to support the general assumption of the dual mode controller, i.e., 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/torAssist}
  \caption{}
  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}

 \noindent
where, $GP$ is the gait phase, $GRF$ is the ground reaction force, and $\delta$ is a threshold value.  This equation explicitly states that an active torque will be applied for each leg during stance and that the legs are passive in their respective swing phases.

Like other control methods presented in this report, the dual mode control method in this section uses the fundamental assumption that it is difficult to measure the interaction forces between the user and hardware system directly as a primary design principle.  To this end, the human-robot interaction force is assumed to be completely represented in terms of the joint torque applied by the operator in the hardware.

To calculate the active torque applied to the joints during the active mode control, the robot dynamics are considered independently from the operator, namely,
\begin{equation}
T_a = M(q)\ddot{q} + C(q,\dot{q})\dot{q}+G(g),
\label{dynamics}
\end{equation}   
where $T_a$ is the joint torque applied by the actuators, $M(q)$ is the inertia matrix, $C(q,\dot{q})$ is the centripetal and Coriolis matrix, and $G(q)$ is the gravitational torques. note that this equation neglects friction as well as other actuation dynamics.

Using the force sensors on the hydraulic actuators, the {\bf human-robot interaction torques} can be approximated as
\begin{equation}
\hat{T}_{hm} = M_n(q)\ddot{q} + C_n(q,\dot{q})\dot{q} + G_n(q) - \hat{T}_a, \notag
\end{equation}
where $M_n$, $C_n$, and $G_n$ are the nominal models of the true inertia, centripetal, and gravitational parameters and $\hat{T}_a$ represents the measured actuator torques.  The $M_n(q)\ddot{q} + C_n(q,\dot{q})\dot{q} + G_n(q) $ portion of $\hat{T}_{hr}$ represents the inverse dynamics component of the closed-loop control system.  The overall joint torque controller, which combines compensation to for operator intent as well as balancing the suit under a gravitational load is then written \[ \tau_\text{active} = K(s) \hat{T}_{hm} + G_n(q),\] where $K(s) = K_p +s K_d,$ i.e., there is closed-loop PD control law operating on the error between the predicted robot dynamics and the measured joint torques.  The block diagram for the active mode controller is shown in Figure \ref{fig:blockDia}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/blockDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:blockDia}   
 \end{figure}
 Note that $G_a$ represents the dynamics of the the system and $G_a'$ the inverse dynamics model in Figure \ref{fig:blockDia}.

In the passive mode, which occurs during the swing phase, the leg of the human-robot system moves quickly and freely under using the passive inertial forces of the overall system.  This is accomplished using bypass valves that connect the pump path to the tank path to achieve mechanical back-drivability.  The foot sensors on each foot are used to trigger the bypass valves and to initiate the pass control mode.

A novel component in the literature of the Hydraulic Lower Extremity Exoskeleton controller is how the transitions between the active and passive control modes are handled.  In particular, an explicit smoothing method is used to address the discontinuity in the command torque at phase transitions.  One solution which is proposed is to use an exponential weighting function of the form
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/weightingForm}
  \caption{}
%  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}

 \noindent
 where $\tau_\text{inv}$ is the inverse dynamics torque associated with the robot model, $f_\text{weighting}$ is the weighting function, $a$ is a sensitivity factor, and $N$ describes the duration of the transition period.  The transition duration period as well as sensitivity factor appear to be hand tuned parameters.

In addition to phase transition smoothing a pre-transition control method which happens at the end of the stance phase prior to the passive swing phase is also used to effect smooth phase transitioning.  The pre-transition control works by measuring the ground reaction force and setting a minimum threshold value.  When the GRF drops below the threshold, the passive swing control mode is entered, i.e., the bypass valve at the actuators are opened.  The objective of this method is to reduce the phase transition delay due to residual pressure in the foot sensors.

>>>>

The Hydraulic Lower Extremity Exoskeleton suit was originally developed for military uses, specifically to allow operators to perform motions without incurring muscle fatigue.  The exoskeleton uses a new {\bf virtual joint torque control} approach, which has been shown to be effective for controlling human robot interaction in situations where the connections between the users and hardware system are relatively unknown or can dramatically change as a function of time.

Additionally, the main hardware design principle is focused on developing a suit that can augment user power and yet move fast when necessary.  The design team selected hydraulic actuators because the weight of electric actuators on leg joints would make the system difficult to move quickly.  Hydraulic actuators have high power density relative to weight and size and generally provide very high force outputs and low impedance.  As a trade-off, hydraulics have reduced accuracy in terms of force control and it is difficult to model their nonlinear characteristics.

@@

The Hydraulic Lower Extremity Exoskeleton's actuators were designed based on human walking data obtained at 4km/h with a 45 kg load.  Based on this data, the hydraulic capacity system was designed to provide 2050psi and 8lpm.  The individual actuators were designed to be double acting at the hip and knee flexion/extension joints. They provide a {\bf hip maximum thrust of 4 kN} and {\bf knee maximum thrust of 4 kN}.  Each joint joint includes three-way servo valves (M 200, Star-Hydraulic) to control rate and direction of bi-directionality, and bypass valves to connect the pump path to tank path for fast, energy-efficient motion during swing phases.

@@

 The Hydraulic Lower Extremity Exoskeleton includes four actively controlled joints in the sagittal plane, with powered knee and hip flexion/extension.  The eight passive joints include one ab/adduction DOF for each hip and 3 DOF for each ankle.

 The hydraulic system is composed of the four actuators and a hydraulic power unit, consisting of a pump, motor, temperature sensor, and various valves.

 On-board sensors measure \textbf{joint angles}, \textbf{actuator forces}, and \textbf{ground reaction forces}.  Three \textbf{absolute angular encoders} are incorporated at the hip, knee, and ankle pitch joints.  In-line \textbf{load cells} are incorporated in each cylinder tube to measure actuator forces, and four single axis load cells are included in foot sensors to be used for gait phase detection.

@@

 \begin{figure}[thpb]
\centering
\includegraphics[width=5.in]{exos/figs/hydLowerExrem/dualModeDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:dualModeDia}   
 \end{figure}
The control design principle for the Hydraulic Lower Extremity Exoskeleton is based on the concept that, for walking gaits, the {\bf stance and swing controllers should be considered separately}.  To this end, the control designers developed an active-passive control method called {\bf Dual-Mode Control}.  Active control is used during stance phases whereas the passive control is used to quickly and freely control the system's legs during swing phases.  Additionally, to handle the inherently discontinuous jumps in the command signals that occur in this dual mode abstraction, a smoothing method is used in conjunction with a pre-transition method that solves for the swing delay due to internal cylinder pressure.  A diagram which highlights the different modes in this control strategy is in Figure \ref{fig:dualModeDia}.

Various components of the dual-mode control strategy, at the hardware level, are presented in Figure \ref{fig:suitDia}.  
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/suitDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:suitDia}   
 \end{figure}
This figure contains a {\bf gait phase estimation block}, which uses the foot sensors to estimate the stance mode of the system, a {\bf human intention block}, which uses angular encoders and joint force sensors to estimate the motion of the operator, a {\bf torque controller}, and a {\bf gravity compensation} block.  Note that these various components of the control system are used to support the general assumption of the dual mode controller, i.e., 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/torAssist}
  \caption{}
  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}
 %
 \noindent
where, $GP$ is the gait phase, $GRF$ is the ground reaction force, and $\delta$ is a threshold value.  This equation explicitly states that an active torque will be applied for each leg during stance and that the legs are passive in their respective swing phases.

Like other controllers presented in this report, the dual mode control method in this section uses the fundamental assumption that it is difficult to measure the interaction forces between the user and hardware system directly as a primary design principle.  To this end, the human-robot interaction force is assumed to be completely represented in terms of the joint torque applied by the operator in the hardware.

To calculate the active torque applied to the joints during the active mode control, the robot dynamics are considered independently from the operator, namely,
\begin{equation}
\vT_a = \mM(\vq)\ddot{\vq} + \mC(\vq,\dot{\vq})\dot{\vq}+\mG(\vq),
\label{dynamics}
\end{equation}   
where $\vT_a$ is the joint torque applied by the actuators, $\mM(\vq)$ is the inertia matrix, $\mC(\vq,\dot{\vq})$ is the centripetal and Coriolis matrix, and $\mG(\vq)$ is the gravitational torques. note that this equation neglects friction as well as other actuation dynamics.

Using the force sensors on the hydraulic actuators, the {\bf human-robot interaction torques} can be approximated as
\begin{equation}
\hat{\vT}_{hm} = \mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) - \hat{\vT}_a, \notag
\end{equation}
where $\mM_n$, $\mC_n$, and $\mG_n$ are the nominal models of the true inertia, centripetal, and gravitational parameters and $\hat{\vT}_a$ represents the measured actuator torques.  The $\mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) $ portion of $\hat{\vT}_{hr}$ represents the inverse dynamics component of the closed-loop control system.  The overall joint torque controller, which combines compensation for operator intent as well as balancing of the suit under a gravitational load, is then written \[ {\bf\tau}_\text{active} = K(s) \hat{\vT}_{hm} + \mG_n(\vq),\] where $K(s) = K_p +s K_d,$ i.e., there is closed-loop PD control law operating on the error between the predicted robot dynamics and the measured joint torques.  The block diagram for the active mode controller is shown in Figure \ref{fig:blockDia}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/blockDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:blockDia}   
 \end{figure}
 Note that $\mG_a$ represents the dynamics of the the system and $\mG_a'$ the inverse dynamics model in Figure \ref{fig:blockDia}.

 In the passive mode, which occurs during the swing phase, the leg of the human-robot system moves quickly and freely under using the passive inertial forces of the overall system.  This is accomplished using bypass valves that connect the pump and tank paths to achieve mechanical back-drivability.  The foot sensors trigger the bypass valves and to initiate the pass control mode.

The Hydraulic Lower Extremity Exoskeleton uses a novel scheme to manage transitions between the active and passive control modes.  In particular, the controller applies an explicit smoothing method to address discontinuities in the command torque at phase transitions.  The implementation proposed uses an exponential weighting function of the form
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/weightingForm}
  \caption{}
%  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}

 \noindent
 where ${\bf\tau}_\text{inv}$ is the inverse dynamics torque associated with the robot model, $f_\text{weighting}$ is the weighting function, $a$ is a sensitivity factor, and $N$ describes the duration of the transition period.  The transition duration period as well as sensitivity factor appear to be hand tuned parameters.

In addition to phase transition smoothing, a pre-transition control method is applied at the end of the stance phase prior to the passive swing phase in order to affect smooth phase transitioning.  The pre-transition control works by measuring the ground reaction force and setting a minimum threshold value.  When the GRF drops below the threshold the system enters a passive swing control mode, i.e., the bypass valve at the actuators are opened.  The objective of this method is to reduce the phase transition delay due to residual pressure in the foot sensors.
